Topological Graph Theory

### 2019-05-12 Topology/Paper/paper.tex
Topology: paper is being revised publicly, using double spaced

\documentclass{article}
\input{../../definitions}
\input{../topology}
\usepackage{thmtools}
\usepackage{thm-restate}
\title{Topological Graph Theory}
\begin{document}
\maketitle
\tableofcontents

\section{Introduction}
Topological graph theory is an entire field within topology and as such this paper is by no means meant to cover all of topological graph theory in any depth. This paper instead will first cover a rather shallow overview of the field (section 2), followed by a more in depth study of graphs theory (section 3). The overview will mainly be focused on giving a thorough understanding of what topological graph theory is as well as to briefly cover the history of the field. After the overview we will spend section 3 building up definitions from graph theory that we will need for the later more rigorous parts of the paper. After this we will dive into Kuratowski's Theorem. We will start by defining $K_5$ and $K_{3,3}$ and showing that they are both nonplanar. Up until this point the paper is written to be accessible by anyone, no mater their level of math background. After we show the nonplanarity of $K_5$ and $K_{3,3}$ we continue onto thoroughly prove Kuratowkski's theorem. The audience for this part is expected to have at least an undergraduate understanding of topology and some exposure to graph theory. After Kuratowski's theorem we finish with a brief discussion of what lies beyond Kuratowski's theorem.

\section{Overview}
\subsection{Graphs}
Before we talk about topological graph theory we must first understand what a graph is.

@@

Consider first a set of points, this may be thought of as just drawing dots on a sheet of paper. Each of these points will be called a vertex. Now we may start drawing lines between vertices. Lines may cross over each other and need not be straight. There is no requirement that all vertices have a line going to it. Each of these lines are called an edge. We will simply insist that no edge connects two vertices and that we do not have multiple edges between the same pair of vertices.

@@

Notice that edge $2$ and edge $3$ both connect vertex $A$ to $B$, as well edges $5$ and $6$ do for $B$ and $C$. This is a strict violation of our definition for a graph. The issue of course then comes to what would one call such a beast as this where, presumably, one is able to have as many connections between any pair of vertices and could even have connections from a vertices to itself.

I am particularly glad that you're paying enough attention to notice that the diagram does not depict a graph. This is what we will refer to as a multigraph. It is worth noting that graphs are a type of multigraph, so anything we show to be true for all multigraphs, is also true for all graphs.

Now to solve this problem we need to make one simple observation about how we walk. If we are to go to island (or vertex) we must also leave that island, unless it is the last island we arrive on. This means, that with the exception of the island we start on and end on, each island must have an even number of bridges connected to it. On the multigraph we would say that we need all vertices but a start and end vertex to have an even number of edges. If we look at the multigraph above, we have four vertices that have an odd number of edges connected to them.

\subsection{History}
%TODO Make how short this is less obvious
The seven bridges of K\"onigsberg problem was solved by Euler %TODO Citation (wikipedia for seven bridges of konigsberg)
in 1736. In mathematics this problem is of great historical significance as it is considered to be the beginning of graph theory as well as a sort of precursor to topology. %TODO Citation

\section{Graph Theory Background}

@@

So if this graph can be drawn without intersection, can any graph be drawn without intersections? If some can and some can't how do we tell which can be drawn and which can not? The answer to this comes in Kuratowski's theorem, however before we can even state this theorem we need to build up a bit of terminology for graph theory.

\subsection{Graph Drawing}
A graph drawing is exactly what it sounds like. We've already seen drawings of graphs like the one for K\"oingsberg and two drawings of $k_4$, this means there may be multiple distinct drawings for the same graph. Now we don't need to be very worried about what defines a drawing, that won't be important to us. Simply think of it as the drawing.

@@

\subsection{Face}
Faces are a bit of an odd property here as they fundamentally are actually properties only of plane drawings and not graphs themselves; however the number of faces, as we will see stays consistent between any plane drawing of a planar graph and as such the number of faces is a property that planar graphs have.

A face is defined as a connected space that contains no edges or vertices and itself is bounded by edges and vertices.

@@

Any graph, connected or not, can be broken into connected components. To do this we simply take a vertex and every other vertex connected to it and call that one component, and then repeat with a vertex not in that component. This sort of breaking apart is nice as often we will prove things about connected graphs that are true about all graphs. For example if all components of a graph are planar then the entire graph must be planar, this will be proven below and it allows us to only deal with connected graphs.

@@

\section{Kuratowski's Theorem}
\begin{restatable}[Kuratowski]{theorem}{THEBIGONE}\label{THEBIGONE}
	A graph $G$ is nonplanar if and only if it contains a subgraph that is a subdivision of $K_{3,3}$ or $K_5$
\end{restatable}

@@

\subsection{$K_5$ and $K_{3,3}$}
Our first step is proving that $K_{3,3}$ and $K_5$ are nonplanar. To do this we are going to use the following theorem.

@@

%TODO cite https://www.ics.uci.edu/~eppstein/junkyard/euler/iedge.html

\begin{proof}
	First consider a connected multigraph with no edges, as this is connected we may only have a single vertex. This produces a single and no edges so we find $$v-e+f=1-0+1=2$$ Now that we know that for 0 edges this rule fits, then either the rule ($v-e+f-2$) is true no mater the number of edges, or there is some number at which point this rule breaks, and $v-e+f\not=2$. If there is a number that breaks this rule, then there must be a lowest number (let's call it $k$) that breaks this rule. For any multigraph with $k$ edges, choose any edge $e$ in the graph.
	\begin{itemize}
		\item If $e$ connects a vertex to itself it is a loop and its removal will result in the loss of one edge and one face (each loop creates a face\footnote{this is due to the Jordan curve theorem}). The resulting multigraph has less than $k$ edges and thus we know that for it \begin{align*}2&=v-e+f \\&= v-(e+1)+(f+1)\end{align*} and as the original graph had one more edge and one more face it too would fit this rule.
		\item If $e$ is not a loop we may perform an edge contraction on it and this will reduce both the number of edges and the number of vertices by one. Again this gives us a multigraph with less than $k$ edges so we know that for it \begin{align*}2&=v-e+f\\&=(v+1)-(e+1)+f\end{align*} and as the original graph had one more edge and one more vertex it to would fit this rule.
	\end{itemize}

	From this we know that there can not possibly be some $k$, and thus all multigraphs, regardless of the number of edges must fit this rule.
\end{proof}
This leads to a nice corollary, that will be helpful when talking about graphs.
\begin{corallary}
	Any plane drawing of any multigraph has the same number of faces.
\end{corallary}
\begin{proof}
	Any planar multigraph can be broken into connected components. Each component of a planar multigraph is a connected planar multigraph and thus any planar drawing of it fits the rule $v-e+f=2$.  Now any plane drawing we make of a planar multigraph will be made of parts that already have a constant number of faces.\footnote{This is not a complete proof, to complete it you simply make an induction on the number of components drawn and realize each time you add a component you add exactly $f_{i}-1$ new faces (if $f_i$ is the number of faces in the $i^{\text{th}}$ component drawn).}

@@

	\scalebox{.9}{$K_5$}
\end{center}
\begin{proof}
	$K_5$ is a graph\footnote{Recall that all graphs are multigraphs, but not all multigraphs are graphs.}, and as such an edge can not be a loop and two vertices can share at most one edge. This means that in any plane drawing of a graph, there must be at least 3 edges bordering every face.
	\begin{itemize}
		\item To only have one edge bordering a face would require that the edge be a loop (which we can't have in a graph).
		\item To only have only two edges bordering a face would require that some pair of vertices share more than one edge (which we can't have in a graph).
	\end{itemize}
	Additionally every edge borders no more than two faces, so from this we know that in any graph $2e \ge 3f$ or $\frac23e\ge f$. Now $K_5$ has 10 edges and 5 vertices, so we know that $$\frac2310=\frac{20}3=6.\bar6\ge f$$ We know that $K_5$ is a connected graph (and thus also a connected multigraph) so if $K_5$ were planar then we would have $$2=v-e+f=5-10+f\le5-10+6.\bar6 = 1.\bar6$$ which is clearly false as $$2\not\le1.\bar6$$ Therefore $K_5$ must not be planar.
\end{proof}

@@

	\scalebox{.9}{$K_{3,3}$}
\end{center}
\begin{proof}
	$K_{3,3}$ is a bipartite graph, meaning that there is some way to break the graph into two collections of vertices, where there are no edges that stay within one of these collections. If we look at the drawing of $K_{3,3}$ above we see that the right and left works as these collections for us. There is no edge that goes between two vertices on the right or two vertices on the left. This means that if we can make a plane drawing of $K_{3,3}$ each face must have at least four edges on its boundary. We know from our proof of $K_5$'s nonplanarity that each face in a graph must have at least three edges on its boundary. If any face were to have three edges then the cycle bounding the face would have exactly three vertices as well. Each pair of vertices within these three vertices would share an edge and therefore there is no way to split them up into two groups such that neither group has an edge within it. Additionally every edge borders no more than two faces, so from this we know that in any bipartite graph $2e \ge 4f$ or $e\ge 2f$.

@@

\subsection{Drawings}
Up to this point the paper has been writing such that anyone with a basic mathematical competency and interest could understand and work through the content. From this point on the paper moves a bit faster and assumes knowledge of basic mathematical notation and more experience with math.

@@

\subsection{Subgraph lemmas}

@@

\subsection{The Jordan Curve Theorem}
\begin{theorem}[Jordan curve theorem] %TODO cite wikipedia https://en.wikipedia.org/wiki/Jordan_curve_theorem#Definitions_and_the_statement_of_the_Jordan_theorem
	Let $C$ be a Jordan curve in the plane $\mathbb R^2$. Then its complement, $\mathbb R^2 \setminus C$, consists of exactly two connected components. One of these components is bounded (the interior) and the other is unbounded (the exterior), and the curve C is the boundary of each component.
\end{theorem}

\begin{definition}[Jordan curve]
	A Jordan curve is a curve, $f:\ccint01\to \mathbb R^2$, with the following properties:
	\begin{itemize}
		\item $f$ is continuous
		\item $f$ is injective on the interval $\coint 01$
		\item $f(0) = f(1)$
	\end{itemize}
\end{definition}
\begin{lemma}\label{JordansLemma}
	For any plane drawing of a graph, $G$, containing cycle $C$, $C$ is a Jordan curve in the drawing.
\end{lemma}
\begin{proof}
%TODO ask Shaw what he thinks
	Consider a plane drawing of a graph $G$ with cycle $C$. We may write out $C$ as $v_1e_1v_2e_2\ldots v_ke_kv_1$. We know that each of these $e_i$s have a corresponding curve $\hat e_i$ that starts at the end of the previous ($\hat e_{i-1}$) curve and ends at the beginning of the next ($\hat e_{i+1}$) curve. This means we can construct a new continuous curve simply by stringing each of these together and then resize the domain. What we end with will then be a curve $f:\ccint 01\to\mathbb R^2$ that is continuous. We will also have $f(0) = f(1)$ as $v_1$ as $e_1$ starts at $v_1$ and $e_k$ ends at $v_1$. We will also have $f$ injective on $\coint{0}{1}$ as we know that edges do not overlap each other (this is formalized in the fourth bullet point of our definition for a drawing). This then means that our cycle creates a Jordan curve.
\end{proof}

@@

\begin{lemma} \label{outside}
	For any face, $F$, in a plane drawing, $D$, of a graph $G$, there is a corresponding drawing plane drawing of $G$, $D'$, where all edges and vertices that border $F$ in $D$, now border the outer face.
\end{lemma}
\begin{proof}
	Let $G$ be a graph with face $F$ in plane drawing $D$. Now let $\inv f$ be a stereographic projection from the punctured sphere onto the plane. Therefore $f(D)$ is a drawing of $G$ on the punctured sphere. If $f(D)$ is a drawing of $G$ on the punctured sphere then it must also be a drawing of $G$ on the sphere. Now we pick a point on the interior of $F$ and puncture the sphere there rather than where our original puncture was.\footnote{There is a problem in this proof. How do we know that $F$ has an interior? Simply put if we have a finite number of vertices and never use any space filling curves for edges we will never fill space. We know that we are never forced to use a space filling curve as for any space filling curve $c$ we may create a new curve $c'$ that connects the same end points, stays within $c$'s image, and is not space filling.} We may now use a sterographic projection, $g$ to map $f(D)$ from the new punctured sphere back to $\mathbb R^2$ creating $D' = g(f(D))$. We know that in $D'$ the face that corresponds to $F$ is the outer face as it contained the punctured point in $\inv g(D')$. %TODO Ask shaw if I should prove that the puctured point in the sphere corresponds to a limit point
\end{proof}

@@

\begin{theorem}[Menger's theorem]\label{menger}
	A non-empty graph is $n$ connected if and only if among all pairs of vertices $a\not=b$ in $G$ the least number of internally disjoint paths from $a$ to $b$ is $n$.
\end{theorem}
%TODO cite Wikipeida menger's theorem
\begin{proof}
	This is trivially true on complete graphs, so for the rest of this proof we will only be considering graphs that are not complete.

@@

\begin{proof}
	First consider the complete graph. We know that $K_4$ is planar (see below), therefore by the contrapositive of lemma \ref{subgraph} all subsets of $K_4$, are planar. Any complete graph $K_i$ with $i < 4$ is a subgraph of $K_4$ (simply remove any $4-i$ vertices from $K_4$) so a nonplanar complete graph must be at least 4 connected.

@@

\begin{lemma}\label{lobes}
	Let $G$ be a nonplanar 2 connected graph with minimal cut set $S=\set{x,y}$. Further if $L$ is a lobe of $S$ let $L^+$ be the graph of $L$ with an edge added between $x$ and $y$. There exists some lobe $L$ of $S$ for which $L^+$ is nonplanar. 
\end{lemma}
%TODO cite http://www.cs.rpi.edu/~goldberg/14-GT/19-kurat.pdf
\begin{proof}
	Let $G$ be a graph with lobes $L_1, L_2, \ldots, L_n$ of separating set $S=\set{x,y}$. Suppose that for all $i$, $L_i^+$ is planar. We may then draw $L_1^+$ on the plane. If we have a planar drawing of $\unionfrom{j = 1}{i-1}{L_j^+}$ we can then embed a drawing of $L_i^+$ where the edge between $x$ and $y$ borders the outer edge into a face that is bordered by the edge between $x$ and $y$ in our planar drawing of $\unionfrom{j = 1}{i-1}{L_j^+}$. We then can redraw this embedding such that $x$ and $y$ in $L_i^+$ are at the same points as $x$ and $y$ in $\unionfrom{j = 1}{i-1}{L_j^+}$ yielding a planar drawing of $\unionfrom{j=1}{i}{L_j^+}$. By induction then $\unionfrom{i=1}{n}{L_i^+}$ is planar. Clearly $G \subset \unionfrom{i=1}{n}{L_i^+}$ so by lemma \ref{subgraph} we know $G$ must be planar.
\end{proof}

\begin{lemma}\label{3connect}
	If $G$ is a graph with the properties listed below, then $G$ is at least 3-connected.
	\begin{enumerate}
		\item $G$ does not contain any subgraph that is contactable to of $K_{3,3}$ or $K_5$.
		\item $G$ is nonplanar.
		\item Of all graphs satisfying properties (1) and (2), $G$ has the least number of edges.
		\item Of all graphs satisfying properties (1), (2) and (3), $G$ has the least number of vertices.
	\end{enumerate}
\end{lemma}
%TODO cite http://www.cs.rpi.edu/~goldberg/14-GT/19-kurat.pdf

\begin{proof}
	Let $G$ be a graph fitting the above properties. Any proper subgraph, $H$, of $G$ will be missing at least one vertex or one edge from $G$. If $H$ is missing a vertex from $G$, then $H$ must break one of the first 3 properties. $H$ can not contain a subgraph that is contactable to $K_{3,3}$ or $K_5$ as then $G$ would also contain it and $H$ can not have more vertices than $G$. This leaves only that $H$ must be planar. If $H$ is missing an edge from $G$, then $H$ must break one of the first 2 properties. Again $H$ can not contain a subgraph that is contactable to $K_{3,3}$ or $k_5$ as then $G$ would also contain it. This again leaves only that $H$ must be planar. This means that all proper subgraphs $H$ of $G$ are planar so $G$ is minimally nonplanar. Additionally we know that $G$ is not a complete graph as all complete graphs either contain $K_5$ or are planar.

@@

\begin{proof}
	Suppose that $G$ is a graph with the following properties.
	\begin{enumerate}
		\item $G$ does not contain any subgraph that is contactable to of $K_{3,3}$ or $K_5$.
		\item $G$ is nonplanar.
		\item Of all graphs satisfying properties (1) and (2), $G$ has the least number of edges.
		\item Of all graphs satisfying properties (1), (2) and (3), $G$ has the least number of vertices.
	\end{enumerate}
	We will try and show that $G$ does not exist.

@@

\section{Beyond Kuratowski's theorem}
Kuratowski's theorem is an unexpectedly elegant statement that says there are only two base case nonplanar graphs, and that all other nonplanar graphs derive their nonplanarity from at least one of these two graphs. A logical next step might be to ask about different surfaces. We've found which graphs can be drawn on a plane, but what about a sphere or a torus? It turns out that the set of graphs that can be drawn on a sphere is the same set of graphs that can be drawn on the plane,\footnote{We actually showed most of what is needed to prove this in our proof for lemma \ref{outside}.} however this is not true for a torus and certainly not true in general. For example $K_{3,3}$ can be drawn on both a torus and a M\"obius strip. So the question naturally arises, do all surfaces have a set of base forbidden graphs like $\set{K_{3,3},K_5}$ is for the plane.

>>>>

\documentclass{article}
\input{../../definitions}
\input{../topology}
\usepackage{thmtools}
\usepackage{thm-restate}
\usepackage{setspace}
\doublespacing
\title{Topological Graph Theory}
\begin{document}
\maketitle
\tableofcontents

\section{Introduction}
Topological graph theory is an entire field within topology and as such this paper is by no means meant to cover all of topological graph theory in any depth. This paper instead will first cover a rather shallow overview of the field (section \ref{SoftOverview}), followed by a more in depth study of graphs theory (section \ref{SoftBackground}). The overview will mainly be focused on giving a thorough understanding of what topological graph theory is as well as to briefly cover the history of the field. After the overview we will spend section \ref{SoftBackground} building up definitions from graph theory that we will need for the later more rigorous parts of the paper. After this we will dive into Kuratowski's Theorem. We will start by defining $K_5$ and $K_{3,3}$ and showing that they are both nonplanar. Up until this point the paper is written to be accessible by anyone, no mater their level of math background. After we show the nonplanarity of $K_5$ and $K_{3,3}$ we continue onto thoroughly prove Kuratowkski's theorem. The audience for this part is expected to have at least an undergraduate understanding of topology and some exposure to graph theory. After Kuratowski's theorem we finish with a brief discussion of what lies beyond Kuratowski's theorem.

\section{Overview} \label{SoftOverview}
\subsection{Graphs}
Before we talk about topological graph theory we must first understand what a graph is.

@@

Consider first a set of points, this may be thought of as just drawing dots on a sheet of paper. Each of these points will be called a vertex. Now we may start drawing lines between vertices. Lines may cross over each other and need not be straight. There is no requirement that all vertices have a line going to it. Each of these lines are called an edge. We will simply insist that no edge connects a vertex to itself and that we do not have multiple edges between the same pair of vertices.

@@

Notice that edge $2$ and edge $3$ both connect vertex $A$ to $B$, as well edges $5$ and $6$ do for $B$ and $C$. This is a strict violation of our definition for a graph. The issue of course then comes to what would one call this graph like object one is able to have as many connections between any pair of vertices and could even have connections from a vertices to itself.

This is what we will call a multigraph. A multigraph is like a graph, but can have an edge that connects a vertex to itself and have two vertices share an edge. It is worth noting that graphs are a type of multigraph, so anything we show to be true for all multigraphs, is also true for all graphs.

Now to solve this problem we need to make one simple observation about how we walk. If we are to go to island (or vertex) we must also leave that island, unless it is the last island we arrive on. This means, that with the exception of the island we start on and end on, each island must have an even number of bridges connected to it. On the multigraph we would say that we need all vertices but a start and end vertex to have an even number of edges. If we look at the multigraph above, we have four vertices that have an odd number of edges connected to them. This means no mater where you start this is no way to take a stroll across every bridge without crossing any bridge twice.

The seven bridges of K\"onigsberg problem was solved by Euler in 1736.\footnote{\url{https://en.wikipedia.org/wiki/Seven_Bridges_of_K\%C3\%B6nigsberg}} In the history of mathematics this problem is of some significance as it is considered to be the beginning of graph theory as well as a sort of precursor to topology.\footnote{\url{https://en.wikipedia.org/wiki/Seven_Bridges_of_K\%C3\%B6nigsberg}}

\subsection{What is topological graph theory?}

Topology the area of math where we deal with different spaces and fundamental concepts like continuity on these spaces. Much of topological graph theory has to do with the relationship between a graph and topological spaces. Once we have a bit more of an overview of graph theory we will dive into Kuratowski's theorem. We use Kuratowski's theorem as a sort of introduction to topological graph theory in this paper as it is a prime example of the relation between a graph and topological spaces.

\section{Graph Theory Background} \label{SoftBackground}

@@

So if this graph can be drawn without intersection, can any graph be drawn without intersections? If some can and some can't how do we tell which can be drawn and which can not? This is the question at the center of Kuratowski's theorem. The rest of this section will serve two purposes. This sections will serve to build up some familiarity with graph theory for those without any previous exposure. The second purpose of this section is to familiarize the reader with definitions and terms from graph theory that will be needed later in our proof of Kuratowski's theorem.

\subsection{Graph Drawing}
A graph drawing is exactly what it sounds like. We've already seen drawings of graphs like the one for K\"oingsberg and two drawings of $k_4$, this means there may be multiple distinct drawings for the same graph. Simply thinking of a drawing as a drawing on a sheet of paper good enough for the vast majority of this paper. We will return to formally define a drawing later in section \ref{Draw} when we need to.

@@

\subsection{Face}
Faces are a bit of an odd property here. Fundamentally faces are a property of plane drawings of a graph, not of the graph itself. A face is defined as a connected space that contains no edges or vertices and itself is bounded by edges and vertices.

@@

Any graph, connected or not, can be broken into connected components. To do this we simply take a vertex and every other vertex connected to it and call that one component, and then repeat with a vertex not in that component. This sort of breaking apart is nice as often we will prove things about connected graphs that are true about all graphs. For example if all components of a graph are planar then the entire graph must be planar, this will be proven in section \ref{SubGraphSec} and it allows us to only deal with connected graphs.

@@

\section{Kuratowski's Theorem}\label{TheTheorem}
\begin{restatable}[Kuratowski]{theorem}{THEBIGONE}\label{THEBIGONE}
	A graph $G$ is nonplanar if and only if it contains a subgraph that is a subdivision of $K_{3,3}$ or $K_5$
\end{restatable}

@@

\subsection{$K_5$ and $K_{3,3}$}\label{k5k33}
Our first step is proving that $K_{3,3}$ and $K_5$ are nonplanar. To do this we are going to use the following theorem.

@@

\begin{proof}\footnote{This proof comes from: \url{https://www.ics.uci.edu/~eppstein/junkyard/euler/iedge.html}.}
	First consider a connected multigraph with no edges, as this is connected we may only have a single vertex. This produces a single and no edges so we find $$v-e+f=1-0+1=2$$ Now that we know that for 0 edges this rule fits, then either the rule ($v-e+f-2$) is true no mater the number of edges, or there is some number at which point this rule breaks, and $v-e+f\not=2$. If there is a number that breaks this rule, then there must be a lowest number (let's call it $k$) that breaks this rule. For any multigraph with $k$ edges, choose any edge $e$ in the graph.
	\begin{itemize}
		\item If $e$ connects a vertex to itself it is a loop and its removal will result in the loss of one edge and one face (each loop creates a face\footnote{this is due to the Jordan curve theorem}). The resulting multigraph has less than $k$ edges and thus we know that for it \begin{align*}2&=v-e+f \\&= v-(e+1)+(f+1)\end{align*} and as the original graph had one more edge and one more face it too would fit this rule.
		\item If $e$ is not a loop we may perform an edge contraction on it and this will reduce both the number of edges and the number of vertices by one. Again this gives us a multigraph with less than $k$ edges so we know that for it \begin{align*}2&=v-e+f\\&=(v+1)-(e+1)+f\end{align*} and as the original graph had one more edge and one more vertex it to would fit this rule.
	\end{itemize}

	From this we know that there can not possibly be some $k$, and thus all multigraphs, regardless of the number of edges must fit this rule.
\end{proof}
This leads to a nice corollary, that will be helpful when talking about graphs.
\begin{corallary}
	Any plane drawing of a specific multigraph $G$ has the same number of faces.
\end{corallary}
\begin{proof}
	Any planar multigraph can be broken into connected components. Each component of a planar multigraph is a connected planar multigraph and thus any planar drawing of it fits the rule $v-e+f=2$.  Now any plane drawing we make of a planar multigraph will be made of parts that already have a constant number of faces.\footnote{This is not a complete proof, to complete it you simply make an induction on the number of components drawn and realize each time you add a component you add exactly $f_{i}-1$ new faces (if $f_i$ is the number of faces in the $i^{\text{th}}$ component drawn).}

@@

	\scalebox{.9}{$K_5$}
\end{center}
\begin{proof}\footnote{This proof comes from \url{http://www.math.cmu.edu/~mradclif/teaching/228F16/Kuratowski.pdf}.}
	$K_5$ is a graph\footnote{Recall that all graphs are multigraphs, but not all multigraphs are graphs.}, and as such an edge can not be a loop and two vertices can share at most one edge. This means that in any plane drawing of a graph, there must be at least 3 edges bordering every face.
	\begin{itemize}
		\item To only have one edge bordering a face would require that the edge be a loop (which we can't have in a graph).
		\item To only have only two edges bordering a face would require that some pair of vertices share more than one edge (which we can't have in a graph).
	\end{itemize}
	Additionally every edge borders no more than two faces, so from this we know that in any graph $2e \ge 3f$ or $\frac23e\ge f$. Now $K_5$ has 10 edges and 5 vertices, so we know that $$\frac2310=\frac{20}3=6.\bar6\ge f$$ We know that $K_5$ is a connected graph (and thus also a connected multigraph) so if $K_5$ were planar then we would have $$2=v-e+f=5-10+f\le5-10+6.\bar6 = 1.\bar6$$ which is clearly false as $$2\not\le1.\bar6$$ Therefore $K_5$ must not be planar.
\end{proof}

@@

	\scalebox{.9}{$K_{3,3}$}
\end{center}
\begin{proof}\footnote{This proof comes from \url{http://www.math.cmu.edu/~mradclif/teaching/228F16/Kuratowski.pdf}.}
	$K_{3,3}$ is a bipartite graph, meaning that there is some way to break the graph into two collections of vertices, where there are no edges that stay within one of these collections. If we look at the drawing of $K_{3,3}$ above we see that the right and left works as these collections for us. There is no edge that goes between two vertices on the right or two vertices on the left. This means that if we can make a plane drawing of $K_{3,3}$ each face must have at least four edges on its boundary. We know from our proof of $K_5$'s nonplanarity that each face in a graph must have at least three edges on its boundary. If any face were to have three edges then the cycle bounding the face would have exactly three vertices as well. Each pair of vertices within these three vertices would share an edge and therefore there is no way to split them up into two groups such that neither group has an edge within it. Additionally every edge borders no more than two faces, so from this we know that in any bipartite graph $2e \ge 4f$ or $e\ge 2f$.

@@

\subsection{Drawings} \label{Draw}
Up to this point the paper has been writing such that anyone with a basic mathematical competency and interest could understand and work through the content. From this point on the paper moves a bit faster and assumes knowledge of basic mathematical notation and more experience with math.

@@

\subsection{Subgraph lemmas} \label{SubGraphSec}

@@

\subsection{The Jordan Curve Theorem}
\begin{theorem}[Jordan curve theorem]
	Let $C$ be a Jordan curve in the plane $\mathbb R^2$. Then its complement, $\mathbb R^2 \setminus C$, consists of exactly two connected components. One of these components is bounded (the interior) and the other is unbounded (the exterior), and the curve C is the boundary of each component.\footnote{This statement of the Jordan curve theorem comes from \url{https://en.wikipedia.org/wiki/Jordan_curve_theorem\#Definitions_and_the_statement_of_the_Jordan_theorem}.}
\end{theorem}

\begin{definition}[Jordan curve]
	A Jordan curve is a curve, $f:\ccint01\to \mathbb R^2$, with the following properties:
	\begin{itemize}
		\item $f$ is continuous
		\item $f$ is injective on the interval $\coint 01$
		\item $f(0) = f(1)$
	\end{itemize}
\end{definition}
\begin{lemma}\label{JordansLemma}
	For any plane drawing of a graph, $G$, containing cycle $C$, $C$ is a Jordan curve in the drawing.
\end{lemma}
\begin{proof}
	Consider a plane drawing of a graph $G$ with cycle $C$. We may write out $C$ as $v_1e_1v_2e_2\ldots v_ke_kv_1$. We know that each of these $e_i$s have a corresponding curve $\hat e_i$ that starts at the end of the previous ($\hat e_{i-1}$) curve and ends at the beginning of the next ($\hat e_{i+1}$) curve. This means we can construct a new continuous curve simply by stringing each of these together and then resize the domain. What we end with will then be a curve $f:\ccint 01\to\mathbb R^2$ that is continuous. We will also have $f(0) = f(1)$ as $v_1$ as $e_1$ starts at $v_1$ and $e_k$ ends at $v_1$. We will also have $f$ injective on $\coint{0}{1}$ as we know that edges do not overlap each other (this is formalized in the fourth bullet point of our definition for a drawing). This then means that our cycle creates a Jordan curve.
\end{proof}

@@

\begin{lemma} \label{outside}
	For any face, $F$, in a plane drawing, $D$, of a graph $G$, there is a corresponding drawing plane drawing of $G$, $D'$, where all edges and vertices that border $F$ in $D$, now border the outer face.
\end{lemma}
\begin{proof}
	Let $G$ be a graph with face $F$ in plane drawing $D$. Now let $\inv f$ be a stereographic projection from the punctured sphere onto the plane. Therefore $f(D)$ is a drawing of $G$ on the punctured sphere. If $f(D)$ is a drawing of $G$ on the punctured sphere then it must also be a drawing of $G$ on the sphere. Now we pick a point on the interior of $F$ and puncture the sphere there rather than where our original puncture was.\footnote{There is a problem in this proof. How do we know that $F$ has an interior? Simply put if we have a finite number of vertices and never use any space filling curves for edges we will never fill space. We know that we are never forced to use a space filling curve as for any space filling curve $c$ we may create a new curve $c'$ that connects the same end points, stays within $c$'s image, and is not space filling.} We may now use a sterographic projection, $g$ to map $f(D)$ from the new punctured sphere back to $\mathbb R^2$ creating $D' = g(f(D))$. We know that in $D'$ the face that corresponds to $F$ is the outer face as it contained the punctured point in $\inv g(D')$.
\end{proof}

@@

\begin{theorem}[Menger's theorem]\label{menger}
	A non-empty graph is $n$ connected if and only if among all pairs of vertices $a\not=b$ in $G$ the least number of internally disjoint paths from $a$ to $b$ is $n$.
\end{theorem}
\begin{proof}
	\footnote{This proof comes from \url{https://en.wikipedia.org/wiki/Menger\%27s_theorem\#Other_proofs}.}
	This is trivially true on complete graphs, so for the rest of this proof we will only be considering graphs that are not complete.

@@

\begin{proof}\footnote{This proof comes from \url{http://math.uchicago.edu/~may/REU2017/REUPapers/Xu,Yifan.pdf}}
	First consider the complete graph. We know that $K_4$ is planar (see below), therefore by lemma \ref{subgraph} all subsets of $K_4$, are planar. Any complete graph $K_i$ with $i < 4$ is a subgraph of $K_4$ (simply remove any $4-i$ vertices from $K_4$) so a nonplanar complete graph must be at least 4 connected.

@@

\begin{lemma}\label{lobes}
	Let $G$ be a nonplanar 2 connected graph with minimal cut set $S=\set{x,y}$. Further if $L$ is a lobe of $S$ let $L^+$ be the graph of $L$ with an edge added between $x$ and $y$. There exists some lobe $L$ of $S$ for which $L^+$ is nonplanar. 
\end{lemma}

\begin{proof} \footnote{This proof comes from \url{http://www.cs.rpi.edu/~goldberg/14-GT/19-kurat.pdf}.}
	Let $G$ be a graph with lobes $L_1, L_2, \ldots, L_n$ of separating set $S=\set{x,y}$. Suppose that for all $i$, $L_i^+$ is planar. We may then draw $L_1^+$ on the plane. If we have a planar drawing of $\unionfrom{j = 1}{i-1}{L_j^+}$ we can then embed a drawing of $L_i^+$ where the edge between $x$ and $y$ borders the outer edge into a face that is bordered by the edge between $x$ and $y$ in our planar drawing of $\unionfrom{j = 1}{i-1}{L_j^+}$. We then can redraw this embedding such that $x$ and $y$ in $L_i^+$ are at the same points as $x$ and $y$ in $\unionfrom{j = 1}{i-1}{L_j^+}$ yielding a planar drawing of $\unionfrom{j=1}{i}{L_j^+}$. By induction then $\unionfrom{i=1}{n}{L_i^+}$ is planar. Clearly $G \subset \unionfrom{i=1}{n}{L_i^+}$ so by lemma \ref{subgraph} we know $G$ must be planar.
\end{proof}

\begin{lemma}\label{3connect}
	If $G$ is a graph with the properties listed below, then $G$ is at least 3-connected.
	\begin{enumerate}
		\item $G$ does not contain any subgraph that is contactable to of $K_{3,3}$ or $K_5$.
		\item $G$ is nonplanar.
		\item Of all graphs satisfying properties (1) and (2), $G$ has the least number of edges.
		\item Of all graphs satisfying properties (1), (2) and (3), $G$ has the least number of vertices.
	\end{enumerate}
\end{lemma}

\begin{proof}\footnote{This proof comes from \url{http://www.cs.rpi.edu/~goldberg/14-GT/19-kurat.pdf}.}
	Let $G$ be a graph fitting the above properties. Any proper subgraph, $H$, of $G$ will be missing at least one vertex or one edge from $G$. If $H$ is missing a vertex from $G$, then $H$ must break one of the first 3 properties. $H$ can not contain a subgraph that is contactable to $K_{3,3}$ or $K_5$ as then $G$ would also contain it and $H$ can not have more vertices than $G$. This leaves only that $H$ must be planar. If $H$ is missing an edge from $G$, then $H$ must break one of the first 2 properties. Again $H$ can not contain a subgraph that is contactable to $K_{3,3}$ or $k_5$ as then $G$ would also contain it. This again leaves only that $H$ must be planar. This means that all proper subgraphs $H$ of $G$ are planar so $G$ is minimally nonplanar. Additionally we know that $G$ is not a complete graph as all complete graphs either contain $K_5$ or are planar.

@@

\begin{proof}\footnote{This proof comes from \url{http://math.uchicago.edu/~may/REU2017/REUPapers/Xu,Yifan.pdf}. This paper also is where most of the structure of everything that has been written from the beginning of this section up to this point comes from.}
	Suppose that $G$ is a graph with the following properties.
	\begin{enumerate}
		\item $G$ does not contain any subgraph that is contactable to of $K_{3,3}$ or $K_5$.
		\item $G$ is nonplanar.
		\item Of all graphs satisfying properties (1) and (2), $G$ has the least number of edges.
		\item Of all graphs satisfying properties (1), (2) and (3), $G$ has the least number of vertices.
	\end{enumerate}
	We will try and show that $G$ does not exist.

@@

\section{Beyond Kuratowski's theorem} \label{Extension}
Kuratowski's theorem is an unexpectedly elegant statement that says there are only two base case nonplanar graphs, and that all other nonplanar graphs derive their nonplanarity from at least one of these two graphs. A logical next step might be to ask about different surfaces. We've found which graphs can be drawn on a plane, but what about a sphere or a torus? It turns out that the set of graphs that can be drawn on a sphere is the same set of graphs that can be drawn on the plane,\footnote{We actually showed most of what is needed to prove this in our proof for lemma \ref{outside}.} however this is not true for a torus and certainly not true in general. For example $K_{3,3}$ can be drawn on both a torus and a M\"obius strip. So the question naturally arises, do all surfaces have a set of base forbidden graphs like $\set{K_{3,3},K_5}$ is for the plane.
